---
title: 'Homework #1: SDGB-7844'
author: "Elaina(Yaogui) Huang"
fordham_id_number: "A18744584"
fordham_email: "yhuang327@fordham.edu"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Each of the exercises below are from our textbook - _Introduction to Scientific Programming and Simulaiton Using R._ The exercises begin on p. 26.  Make sure to comment your code.  Points will be deducted for uncommented code!  

You should create new code blocks if necessary to answer the question.  You should also use plaintext to provide an explanation for your code if the question calls for one.

### Chapter 2 - Exercise 1a. 

```{r}
#Give R assignment statements that set the variable z to the following listed expressions from the textbook
x <- 5
a <-2
b <- 2
(z <- x^(a^b))

```

### Chapter 2 - Exercise 1b. 

```{r}
#assign (x^a)^b to z
(z <- (x^a)^b)

```

### Chapter 2 - Exercise 1c. 

```{r}
#assign 3x^3+2x^2+6x+1(in minimized opration) to z
(z <- (3*x+2)*(x^2+2)-3)
```

### Chapter 2 - Exercise 1d. 

```{r}
#assign the second decimal digit of x into z
x <- 0.456
(z <- as.integer((x*100)%%10))

```

### Chapter 2 - Exercise 1e. 

```{r}
#assign z+1 into z
(z <- z+1)

```


### Chapter 2 - Exercise 2a.

```{r}
#the return should be a sequence from 1-8-1
(x <- c(seq(1,8),seq(7,1)))

```

### Chapter 2 - Exercise 2b.

```{r}
#return repeated sequence
n <- 1 #n start by 1
vector <- c() #create empty vector
while (n < 6) { #start while loop with condition to break when n reach 6
  y<- rep(n, times = n) #assign repetition sequence into y
  n <- n+1 #n increase by 1 in every run until break
  for (p in y) #nested a for loop to assign value
    vector <-c(vector, p) #combine vector's original value with new values attained from the while loop
}
print(vector) #print to show result
```


### Chapter 2 - Exercise 2c. 

```{r}
m <- matrix(1,3,3) #create a 3x3 matrix in m with only #1
diag(m) <- 0 #change the diagonal value to 0
m #print the final result
```

### Chapter 2 - Exercise 2d.

```{r}
A <- matrix(0,3,3)#create 3x3 matrix with only 0s assign to A
m <- c(7,5,3) #create a vector contains 3 numbers assign to m
A[seq(3, by = 2, length = 3)] <- m #assign m to A's extracted value
#extract a sequence from matrix A that is a diagonal 
#A[1,2] <- 2 #change the value in coordination of row 1 column 2 to 2
A #print result
```

### Chapter 2 - Exercise 4

```{r}
vec <- 1:100 #assign numbers from 1 to 100 into vec
(vec_revised <- vec[vec%%2 & vec%%3 & vec%%7]) #use modular division to find out the numbers can't be divided by either 2, 3, or 7
```


### Chapter 2 - Exercise 5a. 

```{r}
queue <- c("Steve", "Russell", "Alison", "Liam")#queue has 4 customers in line
(queue <- c(queue, "Barry")) #now queue has 1 more customer in line
```

### Chapter 2 - Exercise 5b. 

```{r}
served <- "Steve"#Steve is served
(queue <- queue[-which(queue %in% served)])#to match served value in queue by %in% and indicate those matched values index by which() function then use minus sign to eliminate the matched value from queue vector
#assign new vector to replace old values in queue
```

### Chapter 2 - Exercise 5c.

```{r}
(queue <- (append(queue, values = "Pam with 1 item", after = match("Russell",queue)-1)))
#use append function to add value
#assign the new vector into queue
```

### Chapter 2 - Exercise 5d. 

```{r}
impatient <- "Barry" #Barry is impatient
(queue <- queue[-which(queue %in% impatient)])#to match impatient value in queue by %in% and indicate those matched values index by which() function then use minus sign to eliminate the matched value from queue vector
#assign new vector to replace old values in queue
```

### Chapter 2 - Exercise 5e. 

```{r}
impatient <- "Alison" #Alison is impatient
(queue <- queue[-which(queue %in% impatient)])#to match impatient value in queue by %in% and indicate those matched values index by which() function then use minus sign to eliminate the matched value from queue vector
#assign new vector to replace old values in queue

#now look for where Russell is
which(queue=="Russell")
```

### Chapter 2 - Exercise 6

```{r}
rm(list = ls()) 
x <- 1 #successful, x has 1 value
x[3] <- 3#successful, x[3] has 3, x[1:2] are NA NA
y <- c()#successful, y has empty vector, result is NULL
y[2] <- 2 #successful, y[2] has 2
y[3] <- y[1] #successful, result: NA 2 NA
y[2] <- y[4] #successful, result: NA NA NA
#z[1] <- 0#failed, because z never existed, thefore z[1] is not existed
```

### Chapter 2 - Exercise 7

```{r}
diag(5,10,10) #way1 use diag() function to set diagonal as 5 and 10 by 10 row and column
m <- matrix(0,10,10)#way2 create a matrix will full 0s first, then change only the diagonal values
diag(m) <- 5
m
```

